---
title: "Project"
author: "Madison Schultz"
date: "2025-11-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(glmnet)
library(tree)
library(gbm)
library(randomForest)
```

Set up of data
```{r}
housing <- read_csv("Housing.csv")
head(housing)
nrow(housing)
```

Creating Training and Test Data
```{r}
set.seed(123)
test.indices <- sample(1:nrow(housing), 100)
housing.train <- housing[-test.indices,]
housing.test <- housing[test.indices,]

nrow(housing.train)
nrow(housing.test)
```


## Linear Regression
```{r}
lm_model <- lm(price ~ . , data = housing.train)
summary(lm_model)
```


### AIC Model Selection
```{r}
start_model <- lm(price ~ 1, data = housing.train)
step(start_model, direction = "forward", scope = formula(lm_model))
```

```{r}
aic_model <- lm(price ~ . -bedrooms, data = housing.train)
summary(aic_model)
```


### BIC Model Selection
```{r}
step(lm_model, direction = "backward", scope = formula(lm_model))
```
AIC and BIC provide the same model by both taking out bedrooms as a variable in the linear regression

### Calculating Test MSE
```{r}
predict <- predict(aic_model, newdata = housing.test)
test_mse <- mean((housing.test$price - predict)^2)
test_mse
```

## Ridge Regression
```{r}
set.seed(123)
x <- model.matrix(price ~ ., data = housing.train)[,-1]
lambda.list.ridge = 1000 * exp(seq(0, log(1e-5), length = 100))
ridge_mod = cv.glmnet(x, housing.train$price, alpha = 0, lambda = lambda.list.ridge, nfold = 5)

bestlam = ridge_mod$lambda.min

coef(ridge_mod, s = "lambda.min")
```

### Calculating test MSE
```{r}
ridge.pred = predict(ridge_mod, s = bestlam, newx = x)
mean((ridge.pred-housing.train$price)^2)
```
Improves on the test mse but is still very high

## Lasso Regression
```{r}
set.seed(123)
lambda.list.lasso = 2 * exp(seq(0, log(1e-4), length = 100))
lasso_mod = cv.glmnet(x, housing.train$price, alpha = 1, lambda = lambda.list.lasso, nfold = 10)

bestlam = lasso_mod$lambda.min

lasso.coef <- coef(lasso_mod, s = "lambda.min")
lasso.coef
```

### Calculating test mse
```{r}
lasso.pred = predict(lasso_mod, s = bestlam, newx = x)
mean((lasso.pred-housing.train$price)^2)
```
Same test mse as from ridge regression

## Single Decision Tree
```{r}
housing.train[sapply(housing.train, is.character)] <- 
    lapply(housing.train[sapply(housing.train, is.character)], factor)

housing.test[sapply(housing.test, is.character)] <- 
    lapply(housing.test[sapply(housing.test, is.character)], factor)
housing.tree <- tree(price ~ ., data = housing.train)
summary(housing.tree)
```

### Cross Validation to find best tree
```{r}
set.seed(123)
cv <- cv.tree(housing.tree, FUN = prune.tree, K = 5)

best.cv = min(cv$size[cv$dev == min(cv$dev)])
pt.cv = prune.tree(housing.tree, best=best.cv)

pred.pt.cv = predict(pt.cv, housing.test)

mean((pred.pt.cv-housing.test$price)^2)
```
This single decision tree provides a test mse that is the highest.


## Boosting Tree
```{r}
set.seed(123)
boost.housing = gbm(price ~., data = housing.train, n.trees = 1000, shrinkage = 0.01, distribution = "gaussian")

summary(boost.housing)
```

### Calculating Test MSE
```{r}
pred.pt.cv = predict(boost.housing, housing.test)

mean((pred.pt.cv-housing.test$price)^2)
```


## Random Forest
```{r}
set.seed(1)
rf.housing = randomForest(price ~ ., data=housing.train, importance=TRUE)
rf.housing
```


### Calculating Test MSE
```{r}
pred.pt.cv = predict(rf.housing, housing.test)

mean((pred.pt.cv-housing.test$price)^2)
```